target_offtarget

### 2025-01-02 manuscript_figures.Rmd
add notebook for manuscript figures only

## Figure 1: plots of TE experiment data - read counts and read depths

@@

read_count <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(Viral Reads)")

@@

read_count_norm1 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads)")

@@

read_count_norm2 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)")

@@

read_count_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(Viral Reads)")

@@

read_count_norm2_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)")

@@

@@

read_depths <- 
  depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(mean_depth),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12, family = "Helvetica Neue"),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12, family = "Helvetica Neue"),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(mean read depth)")

@@

@@

dedup_per_site <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_site

    TEdata_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
  )

@@

@@

@@

counts_reads_norm<-counts_reads |>
  mutate(norm_counts1 = matched/QC_reads) |>
  mutate(norm_counts2 = matched/QC_reads/length) |>
  mutate(norm_counts3 = matched/length) |>
  mutate(genome_structure = case_when((virus == "Human_adenovirus_40"| virus == "Human_betaherpesvirus") ~ "DNA",
                                      ,.default = "RNA"))

lengths<-counts_reads_norm |>
  select(virus,length) |>
  distinct()

@@

persite_coverage_dedup<-dedup_per_site |>
  group_by(Sample_id,virus,Viral.load,Background,type) |>
  filter(coverage > 0) |>
  summarise(genome_coverage = n()) |>
  left_join(lengths) |>
  mutate(percent_coverage = genome_coverage/length)

persite_coverage_nodedup<-nodedup_per_site |>
  group_by(Sample_id,virus,Viral.load,Background,type) |>
  filter(coverage > 0) |>
  summarise(genome_coverage = n()) |>
  left_join(lengths) |>
  mutate(percent_coverage = genome_coverage/length)

persite_coverage_both<-rbind(persite_coverage_dedup,persite_coverage_nodedup)

coverage_labels<-c("0" = "Control",
                   "100" = "1e+02",
                   "1000" = "1e+03",
                   "100000" = "1e+05",
                   "dedup_TE" = "Deduplicated",
                   "nodedup_TE"= "Non-Deduplicated")

coverage_labels2<-c("0" = "Control",
                    "100" = "1e+02",
                    "1000" = "1e+03",
                    "100000" = "1e+05")

cols2<-c("#BB5566","#004488")

cols4<-c("#DDAA33","#BB5566","#004488")

>>>>

## >>>> Figure 1: plots of TE experiment data - read counts and read depths

@@

read_count <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(Viral Reads)")

@@

read_count_norm1 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts1),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads)")

@@

read_count_norm2 <- counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)")

@@

read_count_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log(matched),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(Viral Reads)")

@@

read_count_norm2_dedup <- 
  counts_reads_norm |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  filter(type == "dedup_TE") |>
  ggplot(aes(
    x = Viral.load,
    y = log(norm_counts2),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_x_log10() +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)")

@@

### --------------------

@@

read_depths <- 
  depths_reads |>
  filter(Background != "p6") |>
  filter(Background != "control") |>
  ggplot(aes(
    x = Viral.load,
    y = log(mean_depth),
    colour = virus
  )) +
  geom_point() +
  geom_smooth(aes(group = virus, linetype = genome_structure), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_few() +
  theme(
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(angle = 0, hjust = 1),
    axis.title.y = element_text(size = 12),
    legend.title = element_blank(),
    legend.position = "top", 
    panel.grid.major = element_line(linewidth=0.1, color = "gray60"),
    panel.grid.minor = element_line(linewidth=0.1, color = "lightgray")
  ) +
  scale_color_manual(
    values = cols,
    labels = c(
      "Human adenovirus 40",
      "Human betaherpesvirus",
      "Human respiratory syncytial virus",
      "Influenza B virus",
      "Mammalian orthoreovirus 3",
      "Zika virus"
    )
  ) +
  scale_x_log10() +
  ylab("Log(mean read depth)")

@@

### --------------------

## Figure S2
```{r}
#metadata

metadata <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/metadata/sampleIDs_TESpikeIn.csv",
    header = TRUE
  )

metadata2 <- metadata %>%
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) %>%
  rename(Sample_id = Sample.ID) %>%
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

counts_dedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_nodedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_reads_norm <- counts_reads %>%
  mutate(norm_counts1 = matched / QC_reads) %>%
  mutate(norm_counts2 = matched / QC_reads / length) %>%
  mutate(norm_counts3 = matched / length) %>%
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  ,
  .default = "RNA"
  ))

####compare library capture pool####

cols3 <- c("#228833", "#AA3377")

metadata3 <- metadata %>%
  select(Sample.ID, Pool.for.sequencing) %>%
  rename(Sample_id = Sample.ID) %>%
  rename(Pool = Pool.for.sequencing)

counts_pool <- left_join(counts_reads_norm, metadata3, by = "Sample_id")

pool_readcounts <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = virus, y = log(matched))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Log(Viral Reads)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_counts_sum <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
                                                                      10)) +
  ylab("Log(Viral Reads)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_readcounts_norm <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = virus, y = log(norm_counts1))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Log(viral reads/cleaned reads)")

pool_norm1_sum <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
                                                                      10)) +
  ylab("Log(viral reads/cleaned reads)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

pool_readcounts_norm2 <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = virus, y = log(norm_counts2))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Log(viral reads/cleaned reads/genome length)")

pool_norm2_sum <- counts_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(matched),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
                                                                      10)) +
  ylab("Log(viral reads/cleaned reads/genome length)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

depth_dedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

depth_nodedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readdepth_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

depths_reads <- left_join(depths_bt_all, metadata2, by = "Sample_id") %>%
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  ,
  .default = "RNA"
  ))

depths_pool <- left_join(depths_reads, metadata3, by = "Sample_id")

pool_readdepths <- depths_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = virus, y = log(mean_depth))) +
  geom_boxplot() +
  facet_grid(Pool ~ type) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  ylab("Log(mean read depth)")

pool_depths_sum <- depths_pool %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = as.character(Viral.load),
    y = log(mean_depth),
    colour = Pool
  )) +
  geom_boxplot() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(axis.title.x = element_blank(), axis.title.y = element_text(size =
                                                                      10)) +
  ylab("Log(mean read depth)") +
  scale_color_manual(values = cols3, labels = c("Pool 1", "Pool 2"))

ggarrange(
  pool_readcounts,
  pool_readcounts_norm,
  pool_readcounts_norm2,
  pool_readdepths,
  nrow = 2,
  ncol = 2,
  heights = c(1, 2),
  common.legend = TRUE
)

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/pool_compare_viruses.pdf")

ggarrange(
  pool_counts_sum,
  pool_norm1_sum,
  pool_norm2_sum,
  pool_depths_sum,
  nrow = 2,
  ncol = 2,
  common.legend = TRUE
)

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/pool_compare_viruses_sum.png",width=10,height=7)

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS2.pdf",width=10,height=7)

### --------------------

## Figure S3
```{r}

##plots of read counts and viral read counts
#November 2024

rm(list = ls())

library(tidyverse)
library(ggpubr)

metadata <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/metadata/sampleIDs_TESpikeIn.csv",
    header = TRUE
  )

metadata2 <- metadata %>%
  select(Sample.ID, Number.of.read.pairs..quality.adaptor.trimmed.) %>%
  rename(Sample_id = Sample.ID) %>%
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

#import read count files

counts_dedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

counts_nodedup_bt <-
  read.table(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv",
    sep = "\t",
    header = TRUE
  )

viral_reads_dedup <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/total_virus_mapped_reads_per_sample_dedup_atcc_ref_20241108.csv",
    header = TRUE
  )

viral_reads_nodedup <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_TE/total_virus_mapped_reads_per_sample_nodedup_atcc_ref_20241108.csv",
    header = TRUE
  )

cols2 <- c("#BB5566", "#004488")

reads_metadata_dedup <-
  left_join(counts_dedup_bt, metadata2, by = "Sample_id")

reads_viral_dedup <-
  left_join(reads_metadata_dedup, viral_reads_dedup, by = "Sample_id")

reads_metadata_nodedup <-
  left_join(counts_nodedup_bt, metadata2, by = "Sample_id")

reads_viral_nodedup <-
  left_join(reads_metadata_nodedup, viral_reads_nodedup, by = "Sample_id")

reads_viral_all <- rbind(reads_viral_dedup, reads_viral_nodedup)

reads_plot <- reads_viral_all %>%
  group_by(Background, Sample_id, Viral.load, type) %>%
  summarise(
    total_reads = (QC_reads * 2),
    viral_reads = total_virus_reads,
    ATCC_reads = sum(matched),
    prop_ATCC = ATCC_reads / total_reads,
    prop_viral = total_virus_reads / total_reads,
    diff = viral_reads - ATCC_reads
  ) %>%
  unique()

total_reads <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = Viral.load,
    y = log(total_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Total Reads") +
  ylab("Log(Total Reads)") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

viral_reads <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = Viral.load,
    y = log(viral_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  theme_bw() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Total Viral Reads") +
  ylab("Log(Total Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

ATCC_reads <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(
    x = Viral.load,
    y = log(ATCC_reads),
    colour = Background
  )) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  theme_bw() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Spike In Viral Reads") +
  ylab("Log(Spike In Viral Reads)") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

prop_ATCC <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = Viral.load, y = prop_ATCC, colour = Background)) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme_bw() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Proportion Spike In Viral") +
  ylab("Spike In Viral Reads/Total Reads") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

prop_viral <- reads_plot %>%
  filter(Background != "p6") %>%
  filter(Background != "control") %>%
  ggplot(aes(x = Viral.load, y = prop_viral, colour = Background)) +
  geom_point() +
  geom_smooth(aes(group = Background), se = FALSE) +
  theme_bw() +
  facet_grid( ~ type, labeller = as_labeller(facet_names)) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 10)
  ) +
  scale_x_log10() +
  ggtitle("Proportion Viral") +
  ylab("Viral Reads/Total Reads") +
  scale_color_manual(values = cols2, labels = c("ME_P1", "ME_P2"))

#ggarrange(total_reads,viral_reads,ATCC_reads,prop_viral,prop_ATCC,nrow=2,ncol=3,common.legend = TRUE)

ggarrange(total_reads,
          viral_reads,
          prop_viral,
          nrow = 3,
          common.legend = TRUE)

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/compare_spike_ins_atcc/backgrounds_reads.png")

#ggsave("/Users/laura/Dropbox/glasgow/github/te_ug_rodents/figures/manuscript_figures_pdf/FigureS3.pdf")

####compare proportion viral reads per pool with shotgun####

metadata3 <- metadata %>%
  select(Sample.ID,
         Number.of.read.pairs..quality.adaptor.trimmed.,
         Pool.for.sequencing) %>%
  rename(Sample_id = Sample.ID) %>%
  rename(QC_reads = Number.of.read.pairs..quality.adaptor.trimmed.)

reads_viral_dedup3 <-
  full_join(metadata3, viral_reads_dedup, by = "Sample_id")

reads_viral_nodedup3 <-
  full_join(metadata3, viral_reads_nodedup, by = "Sample_id")

reads_viral_dedup3 %>%
  group_by(Pool.for.sequencing) %>%
  summarise(
    viral_reads = sum(total_virus_reads),
    total_reads = sum(QC_reads * 2),
    prop_viral = viral_reads / total_reads
  )

polyomics <-
  read.csv(
    "/Users/laura/Dropbox/glasgow/github/te_ug_rodents/data_polyomics/total_virus_mapped_reads_per_sample_dedup.csv"
  )

#read data from polyomics_indexes_stefano document
total_reads <- data.frame(total_reads = c(5942760, 7714275))

keeps <- c("RNA-Msp-p2", "RNA-Msp-p8")

polyomics_read_prop <- polyomics %>%
  filter(Sample_id %in% keeps)

polyomics_read_prop2 <- cbind(polyomics_read_prop, total_reads)

polyomics_read_prop2 %>%
  summarise(prop_viral = total_virus_reads / total_reads)

### --------------------

@@

unzip(zipfile = "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv.zip", exdir = "data_TE/")

unzip(zipfile = "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv.zip", exdir = "data_TE/")

dedup_per_site <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
  )

@@

file.remove("data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_dedup_atcc_ref.tsv")

file.remove( "data_TE/TE_sequencing_experiment_readdepth_per_site_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv")

@@

#import and combine read count files

#deduplicated and non-deduplicated

counts_dedup_bt <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_dedup_atcc_ref.tsv"
  )

counts_nodedup_bt <-
  read_tsv(
    "data_TE/TE_sequencing_experiment_readcount_per_sample_and_virus_bowtie2_nodedup_atcc_ref.tsv"
  )

@@

counts_reads_norm <- 
  counts_reads |>
  mutate(norm_counts1 = matched / QC_reads) |>
  mutate(norm_counts2 = matched / QC_reads / length) |>
  mutate(norm_counts3 = matched / length) |>
  mutate(genome_structure = case_when((
    virus == "Human_adenovirus_40" |
      virus == "Human_betaherpesvirus"
  ) ~ "DNA",
  .default = "RNA"
  ))

lengths <- 
  counts_reads_norm |>
  select(virus, length) |>
  distinct()

@@

persite_coverage_dedup <- 
  dedup_per_site |>
  rename(Viral.load = `Viral load`) |>
  group_by(Sample_id, virus, Viral.load, Background, type) |>
  filter(coverage > 0) |>
  summarise(genome_coverage = n()) |>
  left_join(lengths) |>
  mutate(percent_coverage = genome_coverage / length)

persite_coverage_nodedup <- 
  nodedup_per_site |>
  rename(Viral.load = `Viral load`) |>
  group_by(Sample_id, virus, Viral.load, Background, type) |>
  filter(coverage > 0) |>
  summarise(genome_coverage = n()) |>
  left_join(lengths) |>
  mutate(percent_coverage = genome_coverage / length)

persite_coverage_both <-
  rbind(persite_coverage_dedup, 
        persite_coverage_nodedup)

coverage_labels <- c("0" = "Control",
                     "100" = "10^{2}",
                     "1000" = "10^{3}",
                     "10000" = "10^{5}",
                     "dedup_TE" = "Deduplicated",
                     "nodedup_TE"= "Non-Deduplicated")

coverage_labels2 <- c("0" = "Control",
                      "100" = "10^{2}",
                      "1000" = "10^{3}",
                      "10000" = "10^{5}")
```

### Make plots
```{r, echo = FALSE, warning=FALSE, message=FALSE, fig.height=4, fig.width=6}

cols2 <- c("#BB5566",
           "#004488")

cols4 <- c("#DDAA33",
           "#BB5566",
           "#004488")

persite_coverage_both$Viral.load <- factor(persite_coverage_both$Viral.load, labels = c("0",
    "10^{2}", 
    "10^{3}", 
    "10^{5}"
    ))
